---
output:
  github_document:
    html_preview: true
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

<!-- README.Rmd generates README.md. -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  # collapse = TRUE,
  fig.align = "center",
  comment = "#>",
  fig.path = "man/figures/",
  message = FALSE,
  warning = FALSE
)

options(width = 100)
```


# `{homophily}`

<!-- badges: start -->
[![Lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![GitHub last commit](https://img.shields.io/github/last-commit/knapply/homophily.svg)](https://github.com/knapply/homophily/commits/master)
<!-- [![Codecov test coverage](https://codecov.io/gh/knapply/homophily/branch/master/graph/badge.svg)](https://codecov.io/gh/knapply/homophily?branch=master) -->
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/knapply/homophily?branch=master&svg=true)](https://ci.appveyor.com/project/knapply/homophily)
[![Travis-CI Build Status](https://travis-ci.org/knapply/homophily.svg?branch=master)](https://travis-ci.org/knapply/homophily)
[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
[![Depends](https://img.shields.io/badge/Depends-GNU_R>=3.6-blue.svg)](https://www.r-project.org/)
[![GitHub code size in bytes](https://img.shields.io/github/languages/code-size/knapply/homophily.svg)](https://github.com/knapply/homophily)
[![HitCount](http://hits.dwyl.io/knapply/homophily.svg)](http://hits.dwyl.io/knapply/homophily)
<!-- badges: end -->


<center>
  <h4>
    <i>
      They say that "birds of a feather flock together"... but let's not take their word for it.
    </i>
  </h4>
</center>

<br>

Homophily refers to the tendency of actors to share positive ties with other similar actors. The `{homophily}` package provides flexible, generic routines to measure this phenomenon on objects of class `<igraph>` or `<network>`.
    

# Installation

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("knapply/homophily")
```

# Usage

```{r}
library(homophily)
```

```{r}
# undirected <igraph>
data("jemmah_islamiyah", package = "homophily")
# directed <network>
data("sampson", package = "ergm")
```


## Mixing Matrix

We can easily build classical mixing matrices for undirected and directed graphs.

```{r}
as_mixing_matrix(jemmah_islamiyah, row_attr = "role")
as_mixing_matrix(samplike, row_attr = "group")
```

We can also build _generalized_ mixing matrices to explore homophily beyond the topographic level.

For example, if we want to explore ties between each individual node and a group attribute, we can provide arguments to both `row_attr=` and `col_attr=`.

```{r}
as_mixing_matrix(samplike, row_attr = "vertex.names", col_attr = "group")
```

We can even explore homophily _across_ group attributes.

```{r}
as_mixing_matrix(samplike, row_attr = "cloisterville", col_attr = "group")
```


For directed graphs, the default behavior considers both outgoing and inbound ties, but you can provide `"out"` or `"in"` to `direction=` as desired.

```{r}
as_mixing_matrix(samplike, row_attr = "cloisterville", col_attr = "group",
                 direction = "out")
as_mixing_matrix(samplike, row_attr = "cloisterville", col_attr = "group",
                 direction = "in")
```

## E-I Index

```{r}
ei_index(jemmah_islamiyah, node_attr_name = "role")
ei_index(jemmah_islamiyah, node_attr_name = "role", scope = "group")
ei_index(jemmah_islamiyah, node_attr_name = "role", scope = "node")

ei_index(samplike, node_attr_name = "group")
ei_index(samplike, node_attr_name = "group", scope = "group")
ei_index(samplike, node_attr_name = "group", scope = "node")
```

## Assortativity

```{r}
assortativity_attr(jemmah_islamiyah, node_attr_name = "role")
assortativity_attr(samplike, node_attr_name = "group")
```


# Cite

```{r}
citation("homophily")
```


# `R CMD Check`

```{r}
devtools::check(quiet = TRUE)
```


